Appendix 2: Beta-diversity analysis

### 2021-11-05 code/ms1_beta_diversity.Rmd
updating just in case

---
title: "Beta Diversity through Elevation"
output: html_document
---

# 1. Prepare environment
## 1.1 Packages 
```{r message=FALSE, warning=FALSE}
library(sf)
library(magick)
library(gdm)
library(patchwork)
library(bipartite)
library(vegan)
library(ecodist)
library(flextable)
library(tidyverse)
library(lubridate)
library(cowplot)
```
## 1.2 Define functions
### 1.2.1 `sfc_as_cols`: convert geometry column back to x and y columns
```{r}
# A function for converting a geometry column back to x and y columns
# https://github.com/r-spatial/sf/issues/231
sfc_as_cols <- function(x, names = c("x","y")) {
  stopifnot(inherits(x,"sf") && inherits(sf::st_geometry(x),"sfc_POINT"))
  ret <- sf::st_coordinates(x)
  ret <- tibble::as_tibble(ret)
  stopifnot(length(names) == ncol(ret))
  x <- x[ , !names(x) %in% names]
  ret <- setNames(ret,names)
  dplyr::bind_cols(x,ret)
}
```

### 1.2.2 `web_prep`: convert networks to webs arrays
```{r}
# Function to convert networks to webs arrays
web_prep <- function(net) {

@@

### 1.2.3 `betalinkr_set1`: betalinkr settings and conversion to tibble
```{r}
# Function for betalinkr settings and conversion to tibble()
betalinkr_set1 <- function(x) {
  betalinkr_multi(x, partition.st = TRUE, partition.rr = TRUE, # partition both species turnover and richness/replacement
                  binary = TRUE, # treating beta-diversity as binary is an appropriate complement to our abundance analysis
                  index = "jaccard") %>% # the Jaccard index is best, I think, for binary analysis
  as_tibble() # for convenience
}
```

### 1.2.4 `interaction_gdm_prep`: convert visitation data into site-pair table for GDM analysis
```{r}
# Function for converting visitation data into site-pair table for GDM analysis 
interaction_gdm_prep <- function(x) {

@@

  # Prep site-pairs for GDM analysis
  sitepair <- formatsitepair(spdat, 
                             bioFormat = 1, # i.e. samples as rows, species as columns, site-dates as rows
                             siteColumn = "site", 
                             dist = "jaccard", # I think Jaccard is the right choice
                             XColumn = "x", YColumn = "y", # spatial columns
                             predData = envdat, 
                             abundance = FALSE) %>% # abundance = FALSE means binary
    filter(s1.elev.mean != s2.elev.mean) # this drops within-site comparisons
}
```

### 1.2.5 `biodata_prep_survey`: prepare the floral survey biodata input to gdm's `formatsitepair()` function
```{r}
# Function for preparing the floral survey biodata input to gdm's formatsitepair function
biodata_prep_survey <- function(surv) {
  surv %>%
    filter(flower.cover > 0) %>%
    mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
    group_by(site, plant.sp) %>%
    summarize(cover = sum(flower.cover)) %>%
    mutate(cover = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
    spread(plant.sp, cover) %>%
    replace(is.na(.), 0) %>%
    arrange(site)
}
```

### 1.2.6 `predData_prep_survey`: prepare the floral survey predictor data input to gdm's `formatsitepair()` function
```{r}
# Function for preparing the floral survey predictor data input to gdm's formatsitepair function
predData_prep_survey <- function(surv) {
  surv %>%
    filter(flower.cover > 0) %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    dplyr::select(site, samples) %>%
    left_join(site_data) %>%
    ungroup() %>%
    mutate(mowing = if_else(management == "mowing", 1, 0),
           grasing = if_else(management == "grasing", 1, 0),
           none = if_else(management == "none", 1, 0)) %>%
    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
    distinct() %>%
    arrange(site) 
}
```

### 1.2.7 `biodata_prep_BB`: prepare the visitation biodata input to gdm's `formatsitepair()` function
```{r}
# Function for preparing the visitation biodata input to gdm's formatsitepair function
biodata_prep_BB <- function(net) {
  net %>%
    filter(caste != "male") %>%
    group_by(site, bb.sp) %>%
    summarize(count = n()) %>%
    mutate(count = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
    spread(bb.sp, count) %>%
    replace(is.na(.), 0) %>%
    ungroup()
}
```

### 1.2.8 `predData_prep_BB`: prepare the visitation predictor data input to gdm's `formatsitepair()` function
```{r}
# Function for preparing the visitation predictor data input to gdm's formatsitepair function
predData_prep_BB <- function(net) {
  net %>%
    filter(caste != "male") %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    dplyr::select(site, samples) %>%
    left_join(site_data) %>%
    ungroup() %>%
    mutate(mowing = if_else(management == "mowing", 1, 0),
           grasing = if_else(management == "grasing", 1, 0),
           none = if_else(management == "none", 1, 0)) %>%
    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
    unique() %>%
    arrange(site) 
}
```

### 1.2.9 `format_set`: setting for gdm's `formatsitepair()` function
```{r}
# Function with setting for gdm's formatsitepair function
format_set <- function(bio, pred) {
  formatsitepair(bio, bioFormat = 1, siteColumn = "site", 
                 dist = "jaccard", XColumn = "x", YColumn = "y", 
                 predData = pred, abundance = FALSE) %>%
  filter(s1.elev.mean != s2.elev.mean) # this avoids comparing repeated measures of the same site (taking advantage of the fact that each site has a unique elevation in our dataset)
} 
```

### 1.2.10 `gdm_tab`: tabulate `gdm.varImp()` output 
```{r}
# Function for tabulating gdm.varImp output 
gdm_tab <- function(x, response) {
  x %>%
    as.data.frame() %>%
    rownames_to_column(var = "predictor") %>%
    mutate(model = rep(response, n()),
           predictor = case_when(
             predictor == "elev.mean" ~ "elevation",
             predictor == "samples" ~ "samples",
             predictor == "Geographic" ~ "distance")
             # predictor == "mowing" ~ "mowing",
             # predictor == "grasing" ~ "grazing",
             # predictor == "none" ~ "none")
           ) %>%
    mutate(importance = round(importance, 2),
           model.var.exp = round(model.var.exp, 2)) %>%
    select(Model = model, Variable = predictor, 
           "Importance [%]" = importance, Var.pval = pval, 
           Model.pval = model.pval, "Model.dev.exp [%]" = model.var.exp) %>%
    tibble() %>%
    mutate(Variable = factor(Variable, levels = c("elevation", "distance", "samples"))) %>%
    arrange(Variable)
}
```

### 1.2.11 `isplineExtract2`: Fixes a bug in `isplineExtract()` that causes it to break when non-default number of splines are used
```{r}
##Extracts Ispline data from a gdm model
isplineExtract2 <- function (model){
  ###########################
  #model = gdmOb
  ###########################
  ##error checking
  ##checks to make sure a gdm model is given
  if(!is(model, "gdm")){
    stop("model argument must be of class = 'gdm'.")
  }

@@

### 1.2.12 `plotUncertainty2` Tweaked to call `isplineExtract2` instead of the broken version
```{r}
##########################################################################
plotUncertainty2 <- function(spTable, sampleSites, bsIters, geo=FALSE, splines=NULL, 
                           knots=NULL, splineCol="blue", errCol="grey80", 
                           plot.linewidth=2.0, plot.layout=c(2,2), parallel=FALSE,
                           cores=2){
  ##A function to plot the uncertantiy of each variable from a GDM model
  #################
  #spTable <- sitePairTab          ##the input site-pair table to subsample from
  #sampleSites <- 0.9     ##fraction of sites that should be retained from site pair table 
  #bsIters <- 50       ##the number of times the site-pair table should be sampled
  #geo <- F              ##rather or not the gdm model takes geography into account, see gdm
  #splines <- NULL       ##splines gdm setting, see gdm
  #knots <- NULL         ##knots gdm setting, see gdm
  #splineCol <- "blue"    ##color of the center line
  #errCol <- "grey80"        ##color of the uncertainty polygon
  #plot.linewidth <- 2.0    ##line width of the center line
  #plot.layout <- c(3,3)    ##number of plots per page
  #parallel <- F            ##rather or not the sampling should happen in parallel processing, to speed it up
  #cores <- 6               ##number of cores to if parallel processing
  #################
  ##function breaks and warnings
  ##makes sure that table is a properly formated site-pair table
  if(!is(spTable, "gdmData")){
    warning("The spTable object is not of class 'gdmData'. See the formatsitepair function for help.")
  }
  ##checks to makes sure table is a matrix or table frame
  if(!(is(spTable, "gdmData") | is(spTable, "matrix") | is(spTable, "data.frame"))){
    stop("The spTable object must be of class 'gdmData', 'matrix', or 'data.frame'.")
  }

@@

## 1.3 Load data
```{r}
net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste  != "male") %>%
  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>% 
  mutate(year = factor(year))

survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. (anytime, anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover)

@@

## 1.4 Load images
```{r}
BB <- image_read("../ms_1/figures/phylopic_BB.png")
FL <- image_read("../ms_1/figures/phylopic_FL.png")
BBxFL <- image_read("../ms_1/figures/phylopic_BBxFL.png")
```

# 2. Process data
## 2.1 BB and FL beta-diversity
```{r}
# Bumble bee beta-diversity based on interaction sampling
bb_beta <- net %>%
  group_by(site, bb.sp) %>%
  summarize(count = n()) %>% 
  pivot_wider(names_from = bb.sp, values_from = count) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("site") %>%
  vegdist(method = "jaccard", binary = TRUE) %>% # Jaccard and binary settings make this comparable to the way I'm running betalinkr
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "BB")

@@

# For each site pair, one will be higher and the other lower; thus, when their order is reversed only the sign of the difference changes. Since no two sites have exactly the same elevation, `filter(elev.diff >= 0)` deduplicates the data by dropping the ordering of a given pair of sites (one of the two orderings) that yields a negative elevation difference while retaining the alternative ordering that yields a positive elevation difference. By >= 0, I retain comparisons of a site to itself. I drop (in the ggplot call) within-site comparisons when I plot against elevation difference, and I isolate within-site comparison when I plot against time. In the latter case, it is also necessary to deduplicate, since each pair of site dates occurs in A-B and B-A. For that, I filter (again, in the ggplot call) so that yday.diff > 0, similarly dropping the orderings that yield negative differences. Note that all this deduplication nonsense is not necessary for the betalinkr call because it deduplicates internally.
```

@@

# Add site data
beta_mit_data <- beta_site %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1 = elev.mean, x1 = x, y1 = y, mgmt1 = management) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y, mgmt1, mgmt2 = management) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         mgmt.diff = if_else(mgmt1 == mgmt2, 0, 1)) %>%
  gather(metric, value, -c(site1, site2, elev1, elev2, 
                           elev.diff, mean.elev, 
                           x1, y1, x2, y2, mgmt1, mgmt2, mgmt.diff)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"),
    metric = factor(metric, levels = c("WN", "ST", "OS", "S", "ST.h", "ST.l", 
                                       "ST.lh", "WN.repl", "WN.rich", 
                                       "OS.repl", "OS.rich")))

# Write to csv for archiving
#write_csv(beta_mit_data, "../Dryad_ms1/partitioned_beta_diversity")
```

# 4. Visualize beta-diversity ~ elevation diff
```{r}
# Set palette
p1 <- c("OS" = "#e7298a", "ST" = "#66a61e", "WN" = "black")

@@

# Patchwork plot for comparing species beta-diversity with interaction beta-diversity
beta_plot <- (beta_plot1 / beta_plot2 / beta_plot3) + 
  plot_annotation(tag_levels = c("A", "B", "C"))
# & 
#   theme_classic()

@@

## 5.2 Bumble bees
```{r}
## GDM is not really built for handling categorical variables; when I include 
## management in the models, they effect is weak, and I don't know how to
## interpret the curves. So, I think it is better to drop the mgmt variables
## from the models and just interpret the continuous effects.

# Prep input for GDM
bioData_BB <- biodata_prep_BB(net)
predData_BB <- predData_prep_BB(net) %>%
  select(-c(mowing, grasing, none))
sitepair_BB <- format_set(bioData_BB, predData_BB)

@@

## 5.3 Flora
```{r}
# Prep input for GDM
bioData_survey <- biodata_prep_survey(survey)
predData_survey <- predData_prep_survey(survey) %>%
  select(-c(mowing, grasing, none))
sitepair_survey <- format_set(bioData_survey, predData_survey)

@@

## 5.4 Tabulate `gdm.varImp()` results
### 5.4.1 Interactions
```{r}
# Extract tables
gdm.varImp_WN.tab1 <- gdm.varImp_WN[[1]] %>%
  as.data.frame() %>%
  dplyr::select(val = 1)

@@

### 5.4.2 Bumble bees
```{r}
# Extract tables
gdm.varImp_BB.tab1 <- gdm.varImp_BB[[1]] %>%
  as.data.frame() %>%
  dplyr::select(val = 1)

@@

### 5.4.3 Flora
```{r}
# Extract tables
gdm.varImp_survey.tab1 <- gdm.varImp_survey[[1]] %>%
  as.data.frame() %>%
  dplyr::select(val = 1)

@@

### 5.4.4 Merge and format
```{r}
colourer <- scales::col_numeric(
  palette = c("transparent", "red"),
  domain = c(0, 100))

# Merge and format
gdm.varImp.tab <- bind_rows(gdm.varImp_BB.tab,
                            gdm.varImp_survey.tab, 
                            gdm.varImp_WN.tab
                            ) %>%
  as_grouped_data(group = "Model") %>%
  flextable::as_flextable() %>%
  merge_v(j = c("Model.pval", "Model.dev.exp [%]")) %>%
  valign(j = c("Model.pval", "Model.dev.exp [%]"), valign = "top") %>%
  autofit() %>%
  bg(j = "Importance [%]",
     bg = colourer, part = "body")

@@

## 5.5. Plot GDM splines
```{r}
# Interaction model
gdm_WN_splines1 <- isplineExtract2(gdm_WN)$x %>%
  as.data.frame() %>%
  dplyr::select(elevation = elev.mean)

@@

# Plot
# gdm_plot_ <- ggplot(gdm_splines, aes(elevation, response, linetype = Level)) +
#   geom_line(size = 0.75) +
#   #geom_vline(xintercept = 1500) +
#   geom_rug(data = tree_line, aes(elevation, color = tree_line), inherit.aes = FALSE) +
#   labs(color = "Tree Line", x = "Elevation", y = "Conditional effect") +
#   theme_classic(14) +
#   guides(linetype = "none") +
#   theme(legend.position = c(0.12, 0.85),
#         plot.margin = margin(1, 3, 0.5, 0.5, "cm"),)
# 
# gdm_plot <- ggdraw() +
#   draw_plot(gdm_plot_) +
#   draw_image(BB, scale = 0.11, x = 0.34, y = -0.19) +
#   draw_image(FL, scale = 0.15, x = 0.33, y = 0.15) +
#   draw_image(BBxFL, scale = 0.18, x = 0.38, y = 0.41)

#ggsave("../ms_1/figures/fig7.pdf", width = 7, height = 5)

@@

# 6. MRM analysis (not included in paper)
```{r, warning=FALSE}
survey_sample_diff <- survey %>%
  group_by(site) %>%
  summarize(samples = length(unique(date))) %>%
  remove_rownames() %>%
  column_to_rownames("site") %>%
  vegdist(method = "euclidean", upper = TRUE) %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "sample.diff")

@@

# Making plotUncertainty not suck
```{r}
test <- plotUncertainty3(sitepair_survey, 
                sampleSites = 0.7, 
                bsIters = 100, 
                geo  = FALSE,
                splines = c(6, 6, 6, 6, 6),
                splineCol = "black", 
                plot.layout = c(1,3))

>>>>

---
title: "Appendix 2: Beta-diversity analysis"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
---

# 1. Prepare environment
## 1.1 Packages 
```{r message=FALSE, warning=FALSE}
library(sf)
library(magick)
library(gdm)
library(patchwork)
library(bipartite)
library(vegan)
library(ecodist)
library(flextable)
library(tidyverse)
library(lubridate)
library(cowplot)
```
## 1.2 Define functions
### 1.2.1 `sfc_as_cols`: convert geometry column back to x and y columns
```{r message=FALSE}
# A function for converting a geometry column back to x and y columns
# https://github.com/r-spatial/sf/issues/231
sfc_as_cols <- function(x, names = c("x","y")) {
  stopifnot(inherits(x,"sf") && inherits(sf::st_geometry(x),"sfc_POINT"))
  ret <- sf::st_coordinates(x)
  ret <- tibble::as_tibble(ret)
  stopifnot(length(names) == ncol(ret))
  x <- x[ , !names(x) %in% names]
  ret <- setNames(ret,names)
  dplyr::bind_cols(x,ret)
}
```

### 1.2.2 `web_prep`: convert networks to webs arrays
```{r message=FALSE}
# Function to convert networks to webs arrays
web_prep <- function(net) {

@@

### 1.2.3 `betalinkr_set1`: betalinkr settings and conversion to tibble
```{r message=FALSE}
# Function for betalinkr settings and conversion to tibble()
betalinkr_set1 <- function(x) {
  betalinkr_multi(x, partition.st = TRUE, 
                  partition.rr = TRUE, # partition both species turnover and richness/replacement
                  binary = TRUE, # treating beta-diversity as binary
                  index = "jaccard") %>% # the Jaccard index is best, I think, for binary analysis
  as_tibble() # for convenience
}
```

### 1.2.4 `interaction_gdm_prep`: convert visitation data into site-pair table for GDM analysis
```{r message=FALSE}
# Function for converting visitation data into site-pair table for GDM analysis 
interaction_gdm_prep <- function(x) {

@@

  # Prep site-pairs for GDM analysis
  sitepair <- formatsitepair(spdat, 
                             bioFormat = 1, # i.e. samples = rows, species = columns, site-dates = rows
                             siteColumn = "site", 
                             dist = "jaccard", # I think Jaccard is the right choice
                             XColumn = "x", YColumn = "y", # spatial columns
                             predData = envdat, 
                             abundance = FALSE) %>% # abundance = FALSE means binary
    filter(s1.elev.mean != s2.elev.mean) # this drops within-site comparisons
}
```

### 1.2.5 `biodata_prep_survey`: prepare the floral survey biodata input to gdm's `formatsitepair()` function
```{r message=FALSE}
# Function for preparing the floral survey biodata input to gdm's formatsitepair function
biodata_prep_survey <- function(surv) {
  surv %>%
    filter(flower.cover > 0) %>%
    mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
    group_by(site, plant.sp) %>%
    summarize(cover = sum(flower.cover)) %>%
    mutate(cover = rep(1, n())) %>% # this makes the data binary
    spread(plant.sp, cover) %>%
    replace(is.na(.), 0) %>%
    arrange(site)
}
```

### 1.2.6 `predData_prep_survey`: prepare the floral survey predictor data input to gdm's `formatsitepair()` function
```{r message=FALSE}
# Function for preparing the floral survey predictor data input to gdm's formatsitepair function
predData_prep_survey <- function(surv) {
  surv %>%
    filter(flower.cover > 0) %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    dplyr::select(site, samples) %>%
    left_join(site_data) %>%
    ungroup() %>%
    mutate(mowing = if_else(management == "mowing", 1, 0),
           grasing = if_else(management == "grasing", 1, 0),
           none = if_else(management == "none", 1, 0)) %>%
    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
    distinct() %>%
    arrange(site) 
}
```

### 1.2.7 `biodata_prep_BB`: prepare the visitation biodata input to gdm's `formatsitepair()` function
```{r message=FALSE}
# Function for preparing the visitation biodata input to gdm's formatsitepair function
biodata_prep_BB <- function(net) {
  net %>%
    filter(caste != "male") %>%
    group_by(site, bb.sp) %>%
    summarize(count = n()) %>%
    mutate(count = rep(1, n())) %>% # this makes the data binary
    spread(bb.sp, count) %>%
    replace(is.na(.), 0) %>%
    ungroup()
}
```

### 1.2.8 `predData_prep_BB`: prepare the visitation predictor data input to gdm's `formatsitepair()` function
```{r message=FALSE}
# Function for preparing the visitation predictor data input to gdm's formatsitepair function
predData_prep_BB <- function(net) {
  net %>%
    filter(caste != "male") %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    dplyr::select(site, samples) %>%
    left_join(site_data) %>%
    ungroup() %>%
    mutate(mowing = if_else(management == "mowing", 1, 0),
           grasing = if_else(management == "grasing", 1, 0),
           none = if_else(management == "none", 1, 0)) %>%
    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
    unique() %>%
    arrange(site) 
}
```

### 1.2.9 `format_set`: setting for gdm's `formatsitepair()` function
```{r message=FALSE}
# Function with setting for gdm's formatsitepair function
format_set <- function(bio, pred) {
  formatsitepair(bio, bioFormat = 1, siteColumn = "site", 
                 dist = "jaccard", XColumn = "x", YColumn = "y", 
                 predData = pred, abundance = FALSE) %>%
  filter(s1.elev.mean != s2.elev.mean) # this avoids comparing repeated measures 
                                       # of the same site (taking advantage of the 
                                       # fact that each site has a unique elevation 
                                       # in our dataset)
} 
```

### 1.2.10 `gdm_tab`: tabulate `gdm.varImp()` output 
```{r message=FALSE}
# Function for tabulating gdm.varImp output 
gdm_tab <- function(x, response) {
  x %>%
    as.data.frame() %>%
    rownames_to_column(var = "predictor") %>%
    mutate(model = rep(response, n()),
           predictor = case_when(
             predictor == "elev.mean" ~ "elevation",
             predictor == "samples" ~ "samples",
             predictor == "Geographic" ~ "distance")
           ) %>%
    mutate(importance = round(importance, 2),
           model.var.exp = round(model.var.exp, 2)) %>%
    select(Model = model, Variable = predictor, 
           "Importance [%]" = importance, Var.pval = pval, 
           Model.pval = model.pval, "Model.dev.exp [%]" = model.var.exp) %>%
    tibble() %>%
    mutate(Variable = factor(Variable, levels = c("elevation", "distance", "samples"))) %>%
    arrange(Variable)
}
```

### 1.2.11 `isplineExtract2`: Fixes a bug in `isplineExtract()` that causes it to break when non-default number of splines are used
```{r message=FALSE}
##Extracts Ispline data from a gdm model
isplineExtract2 <- function (model){
  ###########################
  #model = gdmOb
  ###########################
  ##error checking
  ##checks to make sure a gdm model is given
  if(!is(model, "gdm")){
    stop("model argument must be of class = 'gdm'.")
  }

@@

### 1.2.12 `plotUncertainty2` Tweaked to call `isplineExtract2` instead of the broken version
```{r message=FALSE}
##########################################################################
plotUncertainty2 <- function(spTable, sampleSites, bsIters, geo=FALSE, splines=NULL, 
                           knots=NULL, splineCol="blue", errCol="grey80", 
                           plot.linewidth=2.0, plot.layout=c(2,2), parallel=FALSE,
                           cores=2){
  ##A function to plot the uncertantiy of each variable from a GDM model
  #################
  #spTable <- sitePairTab          ##the input site-pair table to subsample from
  #sampleSites <- 0.9     ##fraction of sites that should be retained from site pair table 
  #bsIters <- 50       ##the number of times the site-pair table should be sampled
  #geo <- F              ##rather or not the gdm model takes geography into account, see gdm
  #splines <- NULL       ##splines gdm setting, see gdm
  #knots <- NULL         ##knots gdm setting, see gdm
  #splineCol <- "blue"    ##color of the center line
  #errCol <- "grey80"        ##color of the uncertainty polygon
  #plot.linewidth <- 2.0    ##line width of the center line
  #plot.layout <- c(3,3)    ##number of plots per page
  #parallel <- F            ##rather or not the sampling should happen in parallel processing, to speed it up
  #cores <- 6               ##number of cores to if parallel processing
  #################
  ##function breaks and warnings
  ##makes sure that table is a properly formated site-pair table
  if(!is(spTable, "gdmData")){
    warning("The spTable object is not of class 'gdmData'. See the formatsitepair function for help.")
  }
  ##checks to makes sure table is a matrix or table frame
  if(!(is(spTable, "gdmData") | is(spTable, "matrix") | is(spTable, "data.frame"))){
    stop("The spTable object must be of class 'gdmData', 'matrix', or 'data.frame'.")
  }

@@

## 1.3 Load data
```{r message=FALSE}
net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste  != "male") %>%
  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>% 
  mutate(year = factor(year))

survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. 
  dplyr::select(site, date, plant.sp, flower.cover)

@@

## 1.4 Load images
```{r message=FALSE}
BB <- image_read("../ms_1/figures/phylopic_BB.png")
FL <- image_read("../ms_1/figures/phylopic_FL.png")
BBxFL <- image_read("../ms_1/figures/phylopic_BBxFL.png")
```

# 2. Process data
## 2.1 BB and FL beta-diversity
```{r message=FALSE}
# Bumble bee beta-diversity based on interaction sampling
bb_beta <- net %>%
  group_by(site, bb.sp) %>%
  summarize(count = n()) %>% 
  pivot_wider(names_from = bb.sp, values_from = count) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("site") %>%
  vegdist(method = "jaccard", 
          binary = TRUE) %>% # Jaccard and binary settings make this comparable to the way I'm running betalinkr
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "BB")

@@

# For each site pair, one will be higher and the other lower; thus, when their order is 
# reversed only the sign of the difference changes. Since no two sites have exactly the 
# same elevation, `filter(elev.diff >= 0)` deduplicates the data by dropping the ordering 
# of a given pair of sites (one of the two orderings) that yields a negative elevation 
# difference while retaining the alternative ordering that yields a positive elevation 
# difference. By >= 0, I retain comparisons of a site to itself. I drop (in the ggplot call) 
# within-site comparisons when I plot against elevation difference, and I isolate within-site 
# comparison when I plot against time. In the latter case, it is also necessary to deduplicate, 
# since each pair of site dates occurs in A-B and B-A. For that, I filter (again, in the 
# ggplot call) so that yday.diff > 0, similarly dropping the orderings that yield negative 
# differences. Note that all this deduplication nonsense is not necessary for the betalinkr
# call because it deduplicates internally.
```

@@

# Add site data
beta_mit_data <- beta_site %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1 = elev.mean, x1 = x, y1 = y, mgmt1 = management) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y, mgmt1, mgmt2 = management) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         mgmt.diff = if_else(mgmt1 == mgmt2, 0, 1)) %>%
  gather(metric, value, -c(site1, site2, elev1, elev2, 
                           elev.diff, mean.elev, 
                           x1, y1, x2, y2, mgmt1, mgmt2, mgmt.diff)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"),
    metric = factor(metric, levels = c("WN", "ST", "OS", "S", "ST.h", "ST.l", 
                                       "ST.lh", "WN.repl", "WN.rich", 
                                       "OS.repl", "OS.rich")))
```

# 4. Visualize beta-diversity ~ elevation diff
```{r message=FALSE, warning=FALSE}
# Set palette
p1 <- c("OS" = "#e7298a", "ST" = "#66a61e", "WN" = "black")

@@

# Patchwork plot for comparing species beta-diversity with interaction beta-diversity
beta_plot <- (beta_plot1 / beta_plot2 / beta_plot3) + 
  plot_annotation(tag_levels = c("A", "B", "C"))

@@

## 5.2 Bumble bees
```{r message=FALSE, warning=FALSE}
# Prep input for GDM
bioData_BB <- biodata_prep_BB(net)
predData_BB <- predData_prep_BB(net) %>%
  select(-c(mowing, grasing, none))
sitepair_BB <- format_set(bioData_BB, predData_BB)

@@

## 5.3 Flora
```{r message=FALSE, warning=FALSE}
# Prep input for GDM
bioData_survey <- biodata_prep_survey(survey)
predData_survey <- predData_prep_survey(survey) %>%
  select(-c(mowing, grasing, none))
sitepair_survey <- format_set(bioData_survey, predData_survey)

@@

## 5.4 Tabulate `gdm.varImp()` results
### 5.4.1 Interactions
```{r message=FALSE}
# Extract tables
gdm.varImp_WN.tab1 <- gdm.varImp_WN[[1]] %>%
  as.data.frame() %>%
  dplyr::select(val = 1)

@@

### 5.4.2 Bumble bees
```{r message=FALSE}
# Extract tables
gdm.varImp_BB.tab1 <- gdm.varImp_BB[[1]] %>%
  as.data.frame() %>%
  dplyr::select(val = 1)

@@

### 5.4.3 Flora
```{r message=FALSE}
# Extract tables
gdm.varImp_survey.tab1 <- gdm.varImp_survey[[1]] %>%
  as.data.frame() %>%
  dplyr::select(val = 1)

@@

### 5.4.4 Merge and format
```{r message=FALSE}
colourer <- scales::col_numeric(
  palette = c("transparent", "red"),
  domain = c(0, 100))

# Merge and format
gdm.varImp.tab <- bind_rows(gdm.varImp_BB.tab,
                            gdm.varImp_survey.tab, 
                            gdm.varImp_WN.tab
                            ) %>%
  as_grouped_data(group = "Model") %>%
  flextable::as_flextable() %>%
  merge_v(j = c("Model.pval", "Model.dev.exp [%]")) %>%
  valign(j = c("Model.pval", "Model.dev.exp [%]"), valign = "top") %>%
  autofit() 
# %>%
#   bg(j = "Importance [%]",
#      bg = colourer, part = "body")

gdm.varImp.tab

@@

## 5.5. Plot GDM splines
```{r message=FALSE}
# Interaction model
gdm_WN_splines1 <- isplineExtract2(gdm_WN)$x %>%
  as.data.frame() %>%
  dplyr::select(elevation = elev.mean)

@@

@@

<!-- # 6. MRM analysis (not included in paper) -->
```{r eval=FALSE, warning=FALSE, include=FALSE}
survey_sample_diff <- survey %>%
  group_by(site) %>%
  summarize(samples = length(unique(date))) %>%
  remove_rownames() %>%
  column_to_rownames("site") %>%
  vegdist(method = "euclidean", upper = TRUE) %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "sample.diff")

@@

<!-- # Making plotUncertainty not suck -->
```{r eval=FALSE, include=FALSE}
test <- plotUncertainty3(sitepair_survey, 
                sampleSites = 0.7, 
                bsIters = 100, 
                geo  = FALSE,
                splines = c(6, 6, 6, 6, 6),
                splineCol = "black", 
                plot.layout = c(1,3))
